---
title: "Exploration of IMDB Quotes"
author: "Tobias Machnitzki, Finn Burgemeister"
date: "5 Dezember 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r}
library(ggplot2)
DATAPATH = "C:/Users/darkl/Dropbox/Transfer1/" 
DATAPATH = paste(DATAPATH, "imdb-quotes.csv", sep="")
d = read.csv(DATAPATH, header=F, stringsAsFactors=F, sep="|", quote="\"")
colnames(d) = c("movie", "year", "episode", "actors", "quote")

# data cleaning
d$year = as.numeric(d$year)
d = d[d$year > 1800 & d$year < 2020 & ! is.na(d$year),]
```
First have a look at what we've got:

```{r}
summary(d)
```

Calculate the number of acters for each movie:
```{r}
numOfActors = rep(777,length(d$actors))
for (i in c(0:length(d$actors))){
numOfActors[i] = length(unlist(strsplit(d$actors[i],",")))
}

qplot(numOfActors,geom="histogram",binwidth=1)
```
There seem to be some outliers on the very right side, lets try to visualize them:

```{r}
qplot(numOfActors,geom="histogram", log="y",binwidth=1)
```
So the huge majority of quotes have 2 to 3 actors involved.


Lets Calculate the number of words in each quote:
```{r}
lengthOfQuote = rep(777,length(d$quote))
for (i in c(0:length(d$quote))){
lengthOfQuote[i] = unlist(sapply(gregexpr("\\w+", d$quote[i]), length))
}
lengthOfQuote = data.frame(lengthOfQuote)

qplot(lengthOfQuote,geom="histogram",log="y",binwidth=100)
```
So how do the length correlate with the number of actors?

```{r}
r = cor(numOfActors,lengthOfQuote)
qplot(numOfActors,lengthOfQuote,geom="point")
```
So actually the length of the article has nothing to do with the number of actors taking part.
```{r}
summary(lengthOfQuote)
```

So most quotes have about 31 words but the longest quote seems to be a monologue.
```{r}

quoteWord = character(length(d$quote))
wordOccurences = rep(777,length(d$quote))
for (i in c(0:length(d))){
  quote_length = 0
  for (word in strsplit(d$movie[i], " ")){
    act_length = length(grep(word,strsplit(d$quote[i], " ")))
    if (act_length > quote_length) {
      quote_length = act_length
      act_word = word
    }
    quoteWord[i] = word 
    wordOccurences[i] = quote_length
  }
}
```

